---
title: "Lei dos Grandes Números e Teorema Central do Limite"
date: "23 de junho de 2015"
output: ioslides_presentation
runtime: shiny
bibliography: lgn_tcl.bib
---

# Lei dos Grandes Números



## Definição


Sejam $X_1, X_2, ...$variáveis aleatórias com esperanças finitas e seja $S_n = X_1+X_2+...+X_n$. A sequência ${X_n:n\geq1}$ satisfaz a *Lei dos Grandes Números* se: 

---------------------------------------------
$\frac{S_n}{n}-E(\frac{S_n}{n}) \xrightarrow{\mathbb{P}r}0$
---------------------------------------------


[@Magalhaes2011]

http://vis.supstat.com/2013/04/law-of-large-numbers/


## Exemplo Interativo 

```{r, echo=FALSE}
inputPanel(
  # Number of observations to be used on chart (from 10^1 to 10^8 observations). 
  sliderInput("sample", label = "Quantidade de Amostras (10^n):", min = 1, max = 8, value = 1, step = 1),
  
  # User defined mean to be used on a selected function. On Uniform case, samples are generated between 0 and 2*<this slider value>.
  sliderInput("media", label = "Média Definida:", min = 0, max= 1, value = 0.5, step=0.01),
  
  # Distribution to be used to build the chart. 
  radioButtons("dist", "Selecione:",
             c("Unif." = 1,
               "Binom." = 2,
               "Expon." = 3,
               "Pois."=4,
               "Cauchy" =5))
)

renderPlot({
  # Line bellow sets random seed. If commentend everytime that RStudio build charts they will be different (even with same sample size)
  set.seed(9486)
  
  
  # Number of observations
  n <- 10^as.numeric(input$sample)

  # Mean as defined by the slider
  mu <- as.numeric(input$media)
  
  # Generation of X (following selected distribution and mean)
  x <- switch(input$dist, 
        "1" = runif(n,0,2*mu),
        "2" = rbinom(n,1,mu),
        "3" = rexp(n,rate=1/mu),
        "4" = rpois(n,lambda=mu),
        "5" = rcauchy(n))
    
  # Cumulative sum of X from 1 to n
  s <- cumsum(x)
  
  # Generation of a vector from 1 to n
  d <- 1:n
  
  # For a better visualization, if n=10 then each dot os plotted are "o" else "."  
  ch <- 'o'
  if(n>10)
    ch <- '.'
  
  # Chart plot
  plot(d,s/d,log='x', pch=ch,ylim=c(0,max(s/d)),xlab='X',ylab='Frequency')
  
  # Line showing sample calculated mean
  abline(h=mean(x),col="GREEN")
  
  # Line showing user defined mean
  abline(h=mu,col="RED")
})
```

# Teorema Central do Limite

## Definição

Sejam ${X_n:n\geq1}$ variáveis aleatórias independentes, identicamente distribuidas e com esperança $\mu$ e variância $\sigma^2$, com $0<\sigma^2<\infty$. Então, para $S_n=X_1+X_2+...+X_n$, temos, 

---------------------------------------------
$\dfrac{S_n-n\mu}{{\sqrt{n\sigma^2}}} \xrightarrow{\mathbb{D}} N(0,1)$
---------------------------------------------

[@Magalhaes2011]

https://www.youtube.com/watch?v=epq-dpMJIxs


## Exemplo Interativo

```{r, echo=FALSE}
inputPanel(
  # Number of observations to be used on chart (from 10^1 to 10^8 observations). 
  sliderInput("sample2", label = "Quantidade de Amostras (10^n):", min = 1, max = 8, value = 1, step = 1),
  
  # Number o random variables to calculate Sn.
  sliderInput("xs", label = "Quantidade de v.a.:", min = 1, max= 30, value = 1, step=1),
  
  # Distribution to be used to build the chart.
  radioButtons("dist2", "Selecione:",
             c("Unif." = 1,
               "Binom." = 2,
               "Beta" = 3,
               "Exp." = 4))
)

renderPlot({
  # Line bellow sets random seed, if commentend everytime that RStudio build charts they will be different (even with same sample size)
  set.seed(9486)
  
  # Number of observations
  n <- 10^as.numeric(input$sample2)
  
  # Number o random variables to calculate Sn
  xs <- as.numeric(input$xs)
  
  # Uniform distribution parameters
  unif_a <- 0 
  unif_b <- 1
  
  # Beta distribution parameters
  beta_a <- 3
  beta_b <- 1
  
  # Binomial distribution parameters
  bino_p <- 0.5
  bino_size <- 1
  
  # Exponential distribution parameter
  lambda <- 1
  
  # Calculating mean (as previously selected)
  mu <- switch(input$dist, 
        "1" = (unif_a+unif_b)/2,      # for Uniform distribution
        "2" = n*bino_p,               # for Binomial distribution
        "3" = beta_a/(beta_a+beta_b), # for Beta distribution
        "4" = 1/lambda)               # for Exponential distribution
  
  # Calculating variance (as previously selected)
  sigma <- switch(input$dist2, 
        "1" = ((unif_b-unif_a)^2)/12,                               # for Uniform distribution
        "2" = n*bino_p*(1-bino_p),                                            # for Binomial distribution
        "3" = (beta_a*beta_b)/(beta_a+beta_b+1)*(beta_a+beta_b)^2,  # for Beta distribution
        "4" = 1/lambda^2)                                           # for Exponential distribution
  
  # Definition of cumulative variable for Sn
  x <- 0
  
  # Multiple generation of X (as previously selected)
  for(i in 1:xs){
    d <- switch(input$dist2, 
        "1" = runif(n,unif_a,unif_b),     # Uniform istribution
        "2" = rbinom(n,bino_size,bino_p), # Binomial distribution
        "3" = rbeta(n,beta_a,beta_b),     # Beta distribution
        "4" = rexp(n,rate=1/mu))          # Exponential distribution
    x <- x + d # Cumulative sum of X
  }
  
  # CLT application
  y <- (x-(xs*mu))/sqrt(xs*sigma^2)
  
  # Chart plot
  hist(y, breaks="FD",probability=TRUE)
  
  # Density
  lines(density(y),col="BLUE")
})
```

## Referências
