---
title: "Law of Large Numbers (LLN) and Central Limit Theorem (CTL)"
date: "30th June 2015"
output: slidy_presentation
bibliography: lgn_tcl.bib
---

```{r definitions, echo=FALSE}
### This block contains just definitions used to generate animations in this presentation.


# Predefinitions

## Number of observations
n <- 10^6

## Defined mean (Uniform will be defined between 0 2*mu)
pmu <- 0.8

# Wherever necessary this seed will be used for random generation
Seed <- 23325

# Number of X variables
xs <- 30



## Beta distribution parameters
beta_a <- 3
beta_b <- 1






# Animation Definition
frames <- 100


## Calculating mean (as previously defined)

mu_beta <- beta_a/(beta_a+beta_b) # for Beta distribution
 

## Calculating variance (as previously defined)
sigma_beta <- (beta_a*beta_b)/(beta_a+beta_b+1)*(beta_a+beta_b)^2  # for Beta distribution

# Progress bar plot 
printProgressBar <- function(len,label,pos){
  # Reset chart 1 (progress bar) 
  par(fg=1)
  
  # Print progress bar
  plot(-5, xlim = c(1,len), ylim = c(0, .3), axes = F, xlab = "", ylab = "", main = label)
  axis(1)
  abline(v=pos, lwd=5, col = rgb(0, 0, 255, 255, maxColorValue=255))
  
}

# Layout definition
setLayout <- function(){
  # Layout Definition
  layout(matrix(c(1, rep(2, 5)), 6, 1))
  
  ## Margins
  par(mar=c(4,4,2,1) + 0.1)
}

# Generate an animation illustrating LLN
# Parameters: pType -> type of distribution (see lines 101 to 106); pN -> Number of Observations
# Obseravtion: Plots every single frame of an animation. Some definitions are setted from global variables. 
generateLLNAnimation <-function(pType, pN){
  
  # Distribution parameters

  ## Uniform distribution parameters
  unif_a <- 0 
  unif_b <- 1
  
  ## Binomial distribution parameters
  bino_p <- 0.5
  bino_size <- 1
  
  ## Exponential and Poisson  distribution parameter
  lambda <- 1
  
  
  setLayout()
  # Animation loop
  for (i in 1:frames) {
    
    # Random seed definition
    set.seed(Seed)
    
    # Observations per frame
    ni <- ceiling(pN^(i/frames))
    
    if(ni==0) ni <-1
    
    # Plot progress bar
    printProgressBar(pN,"n",ni)
    
    # Random values generation
    ## Calculating mean
    mu <- switch(pType,
                 "1" = (unif_b-unif_a)/2,
                 "2" = bino_size*bino_p,
                 "3" = 1/lambda,
                 "4" = lambda,
                 "5" = 0.1) # This is just a reference point, since Cauchy have no mean
    ## Generation of X (following selected distribution and mean)
    x <- switch(pType, 
                  "1" = runif(ni,min=unif_a,max=unif_b),
                  "2" = rbinom(ni,size=bino_size,prob=bino_p),
                  "3" = rexp(ni,rate=1/mu),
                  "4" = rpois(ni,lambda=lambda),
                  "5" = rcauchy(ni))
    
    
    
    
    ## Cumulative summary vector
    summ <- cumsum(x)
    
    ## n-divisor vector
    div <- 1:ni
    
    LLN <- summ/div
    # Chart plot
    
    plot(div,LLN,log='x', pch=20,ylim=c(min(LLN),max(LLN)),xlab='S(x)/n',ylab='Frequency',xlim=c(1,pN))
    ## Current sampĺe mean
    abline(h=mean(x),col="GREEN")
    ## Predefined mean
    abline(h=mu,col="RED")
  }
}


# Generate an animation illustrating CLT
# Parameters: pType -> type of distribution (see lines 101 to 106); pN -> Number of Observations
# Obseravtion: Plots every single frame of an animation. Some definitions are setted from global variables. 
generateCLTAnimation <- function(pType, pN){
  
  # Predefinitions
  
  # Number of X variables
  xs <- 30
  
  # Uniform distribution parameters
  unif_a <- 0 
  unif_b <- 1
  
  # Beta distribution parameters
  beta_a <- 3
  beta_b <- 1
  
  # Binomial distribution parameters
  bino_p <- 0.5
  bino_size <- 1
  
  # Exponential distribution parameter
  lambda <- 1
  
  ## Calculating mean (as previously defined)
  mu <- switch(pType, 
               "1" = (unif_a+unif_b)/2,      # for Uniform distribution
               "2" = bino_size*bino_p,       # for Binomial distribution
               "3" = beta_a/(beta_a+beta_b), # for Beta distribution
               "4" = 1/lambda)               # for Exponential distribution
  
  ## Calculating variance (as previously defined)
  sigma <- switch(pType, 
                  "1" = ((unif_b-unif_a)^2)/12,                               # for Uniform distribution
                  "2" = bino_size*bino_p*(1-bino_p),                          # for Binomial distribution
                  "3" = (beta_a*beta_b)/(beta_a+beta_b+1)*(beta_a+beta_b)^2,  # for Beta distribution
                  "4" = 1/lambda^2)                                           # for Exponential distribution
  
  
  
  # Layout Definition
  layout(matrix(c(1, rep(2, 5)), 6, 1))
  
  ## Margins
  par(mar=c(4,4,2,1) + 0.1)
  
  xn<-0
  
  # Animation loop
  for (z in 1:frames) {
    
    
    
    # Random seed definition
    set.seed(Seed)
    
    
    # Plot progress bar
    printProgressBar(xs,"i",xn)
    
    # Current Xi variables
    
    if(xn !=ceiling(xs*z/frames)){
      xn <- ceiling(xs*z/frames)
      # Random values generation
      
      ## Definition of cumulative variable for Sn
      x<-vector(length=n,mode="numeric")
      y <- vector(length=n,mode="numeric") # CLT application variable
      
      ## Multiple generation of X (as previously selected)
      for(k in 1:xn){
        
        d <- switch(pType, 
                    "1" = runif(pN,unif_a,unif_b),     # Uniform istribution
                    "2" = rbinom(pN,bino_size,bino_p), # Binomial distribution
                    "3" = rbeta(pN,beta_a,beta_b),     # Beta distribution
                    "4" = rexp(pN,rate=1/mu))          # Exponential distribution
        x <- x + d # Cumulative sum of X
      }
      
      # CLT application
      y <- (x-(z*mu))/sqrt(z*sigma^2)
    }
      
    
    # Chart plot
    hist(y, breaks="FD",probability=TRUE,main=paste(xn,"variables and",format(n,scientific = FALSE),"observations"),xlab="CLT Aplication")
    
    # Density
    lines(density(y),col="BLUE")
    
  }
  


}
```


Law of Large Numbers (LLN) and Central Limit Theorem (CTL)
==========================================================

# Law of Large Numbers (LLN)

## Definition

Be $X_1, X_2, ...$ random variables with finite variances and be $S_n = X_1+X_2+...+X_n$. The sequence ${X_n:n\geq1}$ satisfies the *Law of Large Numbers* if: 

---------------------------------------------
$\frac{S_n}{n}-E(\frac{S_n}{n}) \xrightarrow{\mathbb{P}r}0$
---------------------------------------------


[@Magalhaes2011]

http://vis.supstat.com/2013/04/law-of-large-numbers/



## Uniform Distribution Example

```{r LLN-Uniform, fig.width=7, fig.height=6, fig.show='animate',echo=FALSE, interval=0.1}
generateLLNAnimation(1,n)
```  

## Binomial Distribution Example

```{r LLN-Binomial, fig.width=7, fig.height=6, fig.show='animate',echo=FALSE, interval=0.1}
generateLLNAnimation(2,n)
```

## Exponential Distribution Example

```{r LLN-Exponential, fig.width=7, fig.height=6, fig.show='animate',echo=FALSE, interval=0.1}
generateLLNAnimation(3,n)
```


## Poisson Distribution Example

```{r LLN-Poisson, fig.width=7, fig.height=6, fig.show='animate',echo=FALSE, interval=0.1}
generateLLNAnimation(4,n)
```
  

## Cauchy Distribution Example

```{r LLN-Cauchy, fig.width=7, fig.height=6, fig.show='animate',echo=FALSE, interval=0.1}
generateLLNAnimation(5,n)
```


# Central Limit Theorem (CLT)

## Definition

Be ${X_n:n\geq1}$ independent random variables, identically distributed and with expected value $\mu$ and variance $\sigma^2$, with $0<\sigma^2<\infty$. For $S_n=X_1+X_2+...+X_n$,  

---------------------------------------------
$\dfrac{S_n-n\mu}{{\sqrt{n\sigma^2}}} \xrightarrow{\mathbb{D}} N(0,1)$
---------------------------------------------

[@Magalhaes2011]

https://www.youtube.com/watch?v=epq-dpMJIxs



## Uniform Distribution Example

```{r CLT-Uniform, fig.width=7, fig.height=6, fig.show='animate',echo=FALSE, interval=0.5}
generateCLTAnimation(1,n)
```

## Binomial Distribution Example

```{r CLT-Binomial, fig.width=7, fig.height=6, fig.show='animate',echo=FALSE, interval=0.5}
generateCLTAnimation(2,n)
```

## Beta Distribution Example

```{r CLT-Beta, fig.width=7, fig.height=6, fig.show='animate',echo=FALSE, interval=0.5}
generateCLTAnimation(3,n)
```

## Exponential Distribution Example

```{r CLT-Exponential, fig.width=7, fig.height=6, fig.show='animate',echo=FALSE, interval=0.5}
generateCLTAnimation(4,n)
```